---
title: "Rough R - Plots for publication"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

This file contains the code for the plots that appear in our paper.

## 1. Random forests

We create a plot that combines the results of the random forest analysis with a boxplot for English and Hungarian.

Loading packages & data.

```{r}
library(tidyverse)
library(ranger)
library(brms)

source("scripts/rough_helper.r")

eng <- read_csv("final_data/english_norms.csv")
hun <- read_csv("final_data/hun_norms_aggr.csv")
eng_rf_mod_all <- readRDS("models/eng_rf_mod_all.rds")
hun_rf_mod_all <- readRDS("models/hun_rf_mod_all.rds")
```

We now extract variable importances for English & Hungarian, and format them.

```{r}
## Get variable importances for best performing consonant models:

hun_v <- hun_rf_mod_all$variable.importance
eng_v <- eng_rf_mod_all$variable.importance

## Rule of thumb threshold following Strobl et al. (2009):

eng_thresh <- abs(min(eng_v))
hun_thresh <- abs(min(hun_v))

## Get the first 10 variable importances:

hun_v_10 <- sort(hun_v, decreasing=T)[1:10]
eng_v_10 <- sort(eng_v, decreasing=T)[1:10]

## Get rid of redundant text & convert the labels to IPA:

names(eng_v_10) <- names(eng_v_10) %>%
  str_remove("all.") %>%
  str_replace('D', 'ð') %>%
  str_replace('Q', 'ə') %>%
  str_replace('V', 'ʌ')
names(hun_v_10) <- names(hun_v_10) %>%
  str_remove("all.") %>%
  str_replace("Ø", "œ") %>%
  str_replace("O", "o") %>%
  str_replace("a", "aː")
```

```{r}
r_col <- "firebrick1"
other_col <- "deepskyblue1"

## Get boxplot info:

eng_r <- boxplot(Rough.M ~ all.r, data = eng, plot = FALSE)
hun_r <- boxplot(roughness ~ all.r, data = hun, plot = FALSE)

cairo_pdf("graphs/random_forest.pdf", 13, 9.5)

## Complex plot layout
layout_M <- matrix(c(1, 1, 1, 1, 2,
                     3, 3, 3, 3, 4),
	nrow = 2, byrow = TRUE)
	
xfac <- 0.1


par(mai = c(1, 0.4, 0.1, 0.2),
	omi = c(0.75, 1, 1.2, 0.95))
layout(layout_M, heights = c(3, 3))
# Variable importance plot, English:
plot(1, 1, type = 'n', xlab = '', ylab = '',
		xaxt = 'n', yaxt = 'n', bty = 'n',
		xlim = c(-0.4, 2.8), ylim = c(0, 11))

text(x = 1.58, y = 15, cex = 3.8, font = 2, xpd = NA,
	labels = 'Random forest analysis')

abline(h = 1:9, lty = 2, col = 'darkgrey')
abline(h = 10, lty = 2, col = r_col, lwd = 2)

points(x = rev(eng_v_10)[1:9], y = 1:9, pch = 15, cex = 1.5)
points(x = rev(eng_v_10)[10], y = 10, pch = 15, cex = 2.5, col = r_col)

axis(side = 1, at = seq(-0.4, 2.8, 0.4), lwd = 2, cex.axis = 1.5, font = 2)
axis(side = 2, at = 1:10, labels = rev(names(eng_v_10)),
		lwd = 2, cex.axis = 1.8, font = 2, las = 2, family = 'Times New Roman')
mtext(side = 2, text = 'Phonemes', line = 4, font = 2, cex = 1.7)

text(x = -0.6, y = 12.5, xpd = NA, font = 2, cex = 3,
	labels = '(a) English', adj = c(0, 1), pos = 4)

abline(v = 0, lwd = 2, lty = 2)
abline(v = eng_thresh, lty = 1, lwd = 2)
box(lwd = 2)

# Eng boxplot
plot(1, 1,
	type = 'n', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
	ylim = c(-7, 7), xlim = c(0.8, 3.2))
rect(xleft = 1 + xfac, xright = 2 - xfac,
	ybottom = eng_r$stats[2, 1],
	ytop = eng_r$stats[4, 1], col = other_col)
segments(y0 = eng_r$stats[3, 1],
	x0 = 1 + xfac, x1 = 2 - xfac)
segments(x0 = 1.5, y0 = eng_r$stats[4, 1],	# UB
	y1 = eng_r$stats[5, 1])
segments(x0 = 1.5, y0 = eng_r$stats[2, 1],	# LB
	y1 = eng_r$stats[1, 1])
rect(xleft = 2 + xfac, xright = 3 - xfac,
	ybottom = eng_r$stats[2, 2],
	ytop = eng_r$stats[4, 2], col = r_col)
segments(y0 = eng_r$stats[3, 2],
	x0 = 2 + xfac, x1 = 3 - xfac)
segments(x0 = 2.5, y0 = eng_r$stats[4, 2],	# UB
	y1 = eng_r$stats[5, 2])
segments(x0 = 2.5, y0 = eng_r$stats[2, 2],	# LB
	y1 = eng_r$stats[1, 2])
axis(side = 1, at = c(1.5, 2.5),
	labels = c('no /r/', '/r/'), font = 2,
	cex.axis = 1.8)
axis(side = 4, at = c(-7, -3.5, 0, 3.5, 7), font = 2,
	cex.axis = 1.5, las=2)
box(lwd = 2)
mtext(side = 4, text = 'Roughness ratings', line = 4.9, font = 2, cex = 1.7)

# Hungarian:
# Variable importance plot, Hungarian:
plot(1, 1, type = 'n', xlab = '', ylab = '',
		xaxt = 'n', yaxt = 'n', bty = 'n',
		xlim = c(-0.4, 1.6), ylim = c(0, 11))

abline(h = c(1:8, 10), lty = 2, col = 'darkgrey')
abline(h = 9, lty = 2, col = r_col, lwd = 2)

points(x = rev(hun_v_10)[c(1:8, 10)], y = c(1:8, 10), pch = 15, cex = 1.5)
points(x = rev(hun_v_10)[9], y = 9, pch = 15, cex = 2.5, col = r_col)

axis(side = 1, at = seq(-0.4, 1.6, 0.4), lwd = 2, cex.axis = 1.5, font = 2)
axis(side = 2, at = 1:10, labels = rev(names(hun_v_10)),
		lwd = 2, cex.axis = 1.8, font = 2, las = 2, family = 'Times New Roman')
mtext(side = 2, text = 'Phonemes', line = 4, font = 2, cex = 1.7)

text(x = -0.6, y = 12.5, xpd = NA, font = 2, cex = 3,
	labels = str_c('(b) Hungarian'), adj = c(0, 1), pos = 4)

mtext(side = 1, text = 'Relative Variable Importance',
	line = 4, font = 2, cex = 2)
mtext(side = 1, text = '(permutation based)',
	line = 6.4, font = 2, cex = 1.5)

abline(v = 0, lwd = 2, lty = 2)
abline(v = hun_thresh, lty = 1, lwd = 2)
box(lwd = 2)

abline(v = hun_thresh, lty = 2)

# Hun boxplot
plot(1, 1,
	type = 'n', xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
	ylim = c(-7, 7), xlim = c(0.8, 3.2))
rect(xleft = 1 + xfac, xright = 2 - xfac,
	ybottom = hun_r$stats[2, 1],
	ytop = hun_r$stats[4, 1], col = other_col)
segments(y0 = hun_r$stats[3, 1],
	x0 = 1 + xfac, x1 = 2 - xfac)
segments(x0 = 1.5, y0 = hun_r$stats[4, 1],	# UB
	y1 = hun_r$stats[5, 1])
segments(x0 = 1.5, y0 = hun_r$stats[2, 1],	# LB
	y1 = hun_r$stats[1, 1])
rect(xleft = 2 + xfac, xright = 3 - xfac,
	ybottom = hun_r$stats[2, 2],
	ytop = hun_r$stats[4, 2], col = r_col)
segments(x0 = 2.5, y0 = hun_r$stats[4, 2],	# UB
	y1 = hun_r$stats[5, 2])
segments(x0 = 2.5, y0 = hun_r$stats[2, 2],	# LB
	y1 = hun_r$stats[1, 2])
segments(y0 = hun_r$stats[3, 2],
	x0 = 2 + xfac, x1 = 3 - xfac)
box(lwd = 2)
axis(side = 1, at = c(1.5, 2.5),
	labels = c('no /r/', '/r/'), font = 2,
	cex.axis = 1.8)
axis(side = 4, at = c(-7, -3.5, 0, 3.5, 7), font = 2,
	cex.axis = 1.5, las=2)
mtext(side = 4, text = 'Roughness ratings', line = 4.9, font = 2, cex = 1.7)
dev.off()
```

# 2. IE analysis

Loading data, packages, setting up ggplot theme (to match previous graph).

```{r}
library(tidyverse)
library(brms)
library(sf)
library(ggpubr)
library(ggbeeswarm)
library(rnaturalearth)
library(rnaturalearthdata)



source("scripts/rough_helper.r")

trs_IE <- read_csv("final_data/google_translate.csv") %>%
  filter(stock=="Indo-European")
trs_IE$rough.f <- factor(ifelse(trs_IE$rough, "rough words", "smooth words"))

trs_IE_brm_r_mod <- readRDS("models/trs_IE_brm_r_mod.rds")

# now we calculate proportions of /r/ for rough / smooth in each language

trs_IE_prop <- trs_IE %>%
  group_by(language, rough.f) %>%
  summarise(r_prop = mean(r),
            longitude=longitude[1],
            latitude=latitude[1],
            n=n())

theme_rough <- 
theme_classic() + 
  theme(legend.position = "none",
        legend.key.height = unit(2,"line"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.background = element_rect(fill = "transparent"),
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent"),
        strip.text.y = element_text(size = 12, hjust = 0),
        axis.text.x = element_text(size = 12, colour="black", face="bold"),
        axis.text.y = element_text(size = 12, colour="black"),
        axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.margin = unit(c(0.4,0.4,0.4,0.4),"cm"))
```

Create map plot.

```{r ie map plots}
r_col <- "firebrick1"
other_col <- "deepskyblue1"

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

png("graphs/ie_map.png", width=9, height=3.5)
ggplot(world) +
  #geom_polygon(fill = "lightgray", colour = "white", size = 0.2) +
  geom_sf(size=0.2, col="white", fill="lightgrey") +
  coord_sf(xlim = c(-25, 60), ylim = c(30, 70), expand=T, datum=NA) +
  geom_point(data = trs_IE_prop, aes(x = longitude, y = latitude, fill = r_prop, group = NA),
             alpha = 0.9, position = position_jitter(0.1), size = 3,
             shape = 21, color = "black", stroke = 0.3) +
  facet_grid(~rough.f) +
  scale_fill_gradient2(expand=expansion(mult=c(0,0)),
                        name="percentage\nof /r/",
                        low = other_col, high = r_col, midpoint = 0.5,
                        breaks=c(0, 0.25, 0.5, 0.75, 1),
                        labels=c("0%", "25%", "50%", "75%", "100%")) +
  labs(title = "(A) /r/ in rough and smooth words in Indo-European languages",
       subtitle= "(Sinhala and Afrikaans not shown)") +
  theme_rough +
  theme(
    legend.position = "right",
    legend.margin=margin(t=-0.95, b=-0.5, unit="inch"),
    legend.key.height=unit(0.49, "inch"),
    legend.title=element_text(size=12, face="bold", hjust=0),
    legend.text.align=0.5,
    legend.text=element_text(margin=margin(0,0,0,0.05, unit="inch")),
    axis.title = element_blank(),
    axis.text = element_blank(),
    strip.text.x = element_text(vjust=2),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = margin(0.4, -0.15, 0.4, 0.4, unit="cm"))

ggsave("graphs/ie_map.png", width=8.3, height=3.5, dpi=300)
```

Create model prediction plot.

```{r}
# predictions
mod_preds <- logistic_summary(trs_IE_brm_r_mod, dat=trs_IE, outcome="/r/", roughpred="rough", pp_over_zero=T)

mod_dat <- data.frame(rough.f=factor(c("rough words", "smooth words")),
                      pred=c(mean(mod_preds$pred_prob_rough),mean(mod_preds$pred_prob_smooth))*100,
                      ll=c(quantile(mod_preds$pred_prob_rough, 0.025),quantile(mod_preds$pred_prob_smooth, 0.025))*100,
                      ul=c(quantile(mod_preds$pred_prob_rough, 0.975),quantile(mod_preds$pred_prob_smooth, 0.975))*100
)


ggplot(data = trs_IE_prop, aes(x = rough.f, y = r_prop*100)) +
  geom_point(data=trs_IE_prop, 
             aes(fill = r_prop, size = n), alpha = 0.6,
             position = position_quasirandom(width = 0.2),
             shape = 21, color = "black", stroke = 0.3) +
  geom_errorbar(data=mod_dat,
                aes(ymin = ll, ymax = ul, y=NULL), 
                width = 0.075) +
  geom_point(data=mod_dat, 
             aes(y=pred), fill="grey", 
             size = 6, shape = 21) +
  scale_fill_gradient2(guide = F,
                       low=other_col, high=r_col, midpoint=0.5) +
  scale_y_continuous(breaks=seq(0,100,25), labels=paste0(seq(0,100,25), "%"),
                     expand=expansion(mult=0.02)) +
  scale_x_discrete(position = "top") +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "(B) Proportions per language and\nmodel predictions",
       y = "percentage of forms with /r/", 
       x = "\n") +
  theme_rough + 
  theme(plot.margin = unit(c(-0.56,0.4,0.315,0),"cm"),
        plot.title = element_text(size = 14, face = "bold", vjust=-8.1, hjust=1),
        axis.line.y=element_line(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank())
        #axis.ticks.y=element_blank())
ggsave("graphs/ie_model_preds.png", width=3.5, height=3.5, dpi=300)

```

# 3. Cross-linguistic analysis

Loading data, packages, setting up ggplot theme (to match previous graph).

```{r}
library(tidyverse)
library(brms)
library(sf)
library(ggpubr)
library(ggbeeswarm)
library(rnaturalearth)
library(rnaturalearthdata)

source("scripts/rough_helper.r")

xling <- read_csv("final_data/cross_linguistic.csv") %>%
  filter(Meaning %in% c("rough","smooth"),
         Family!="Indo-European")
length(unique(xling$Language))
xling$Meaning <- paste0("‘", xling$Meaning, "’")
xling$Meaning.f <- factor(xling$Meaning)#, levels=c("‘smooth’", "‘rough’"))
# only Google TRS data for Basque:
xling <- filter(xling, !(Language=="Basque" & Dataset=="IDS"))



# some languages happen to have more than one rough / smooth words (very small minority!)
# reduce to single data point via majority rule:
xling_single <- xling %>%
  filter(Trill=="yes") %>%
  group_by(Language,Meaning.f) %>%
  summarise(r = as.logical(round(mean(r))),
            Longitude=Longitude[1],
            Latitude=Latitude[1]) %>%
  ungroup()

xling_brm_rs_logistic_mod_r <- readRDS("models/xling_brm_rs_logistic_mod_r.rds")
xling_nt_brm_rs_logistic_mod_r <- readRDS("models/xling_nt_brm_rs_logistic_mod_r.rds")

theme_rough <- 
theme_classic() + 
  theme(legend.position = "none",
        legend.key.height = unit(2,"line"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        legend.background = element_rect(fill = "transparent"),
        strip.background = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(2, "lines"),
        panel.border = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent"),
        strip.text.y = element_text(size = 12, hjust = 0),
        axis.text.x = element_text(size = 12, colour="black", face="bold"),
        axis.text.y = element_text(size = 12, colour="black"),
        axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = element_text(size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        plot.margin = unit(c(0.4,0.4,0.4,0.4),"cm"))
```

Create map plot.

```{r map plots}
r_col <- "firebrick1"
other_col <- "deepskyblue1"

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

#cairo_pdf("graphs/xling_map.pdf", width=9, height=3)
ggplot(world) +
  geom_sf(size=0.2, col="white", fill="lightgrey") +
  coord_sf(xlim = c(-180, 180), ylim = c(-60, 80), datum=NA) +
  geom_point(data = xling_single, aes(x = Longitude, y = Latitude, fill = r, group = NA),
             alpha = 0.8, position = position_jitter(0.1), size = 2.5,
             col="black", pch=21, stroke=0.1) +
  facet_grid(~Meaning.f) +
  scale_fill_manual(breaks = c(F,T), values = c(other_col, r_col), labels=c("no /r/", "contains /r/")) +
  labs(title = "Trilled /r/ in the words ‘rough’ and ‘smooth’ cross-linguistically") +
  theme_rough +
  theme(
    legend.position = "bottom",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0.1,0.4,-0.4,0.4),"cm"),
    panel.spacing = unit(0, "lines"))
ggsave("graphs/xling_map.png", width=9, height=3)
#dev.off()

```

Create model prediction plot.

```{r}
# predictions
mod_preds_trill <- logistic_summary(xling_brm_rs_logistic_mod_r, dat=xling, outcome="/r/", roughpred="rough", pp_over_zero=T)

mod_dat_trill <- data.frame(Meaning.f=factor(c("‘smooth’", "‘rough’"), levels=c("‘smooth’", "‘rough’")),
                      pred=c(mean(mod_preds_trill$pred_prob_smooth),mean(mod_preds_trill$pred_prob_rough))*100,
                      ll=c(quantile(mod_preds_trill$pred_prob_smooth, 0.025),quantile(mod_preds_trill$pred_prob_rough, 0.025))*100,
                      ul=c(quantile(mod_preds_trill$pred_prob_smooth, 0.975),quantile(mod_preds_trill$pred_prob_rough, 0.975))*100
)

mod_preds_nt <- logistic_summary(xling_nt_brm_rs_logistic_mod_r, dat=xling, outcome="/r/", roughpred="rough", pp_over_zero=T)

mod_dat_nt <- data.frame(Meaning.f=factor(c("‘smooth’", "‘rough’"), levels=c("‘smooth’", "‘rough’")),
                      pred=c(mean(mod_preds_nt$pred_prob_smooth),mean(mod_preds_nt$pred_prob_rough))*100,
                      ll=c(quantile(mod_preds_nt$pred_prob_smooth, 0.025),quantile(mod_preds_nt$pred_prob_rough, 0.025))*100,
                      ul=c(quantile(mod_preds_nt$pred_prob_smooth, 0.975),quantile(mod_preds_nt$pred_prob_rough, 0.975))*100
)

mod_dat <- bind_rows(mod_dat_trill, mod_dat_nt) %>%
  mutate(Trill=factor(c("trilled /r/","trilled /r/","other /r/","other /r/"), 
                      levels=c("trilled /r/","other /r/"))
  )

# raw figures / language family
r_per_family <- xling %>%
  group_by(Family, Trill, Meaning.f) %>%
  summarise(perc_r = mean(r)*100, 
            n=length(unique(Language))) %>%
  ungroup() %>%
  mutate(Trill=recode(Trill, yes="trilled /r/", no="other /r/"),
         Trill=factor(Trill, levels=c("trilled /r/","other /r/"))
  )

ggplot(data = r_per_family, aes(x = Meaning.f, y = perc_r)) +
  facet_grid(. ~ Trill) +
  geom_point(data=r_per_family, 
             aes(color = perc_r, size = n), alpha = 0.5,
             position = position_quasirandom(width = 0.3)) +
  geom_errorbar(data=mod_dat,
                aes(ymin = ll, ymax = ul, y=NULL), 
                width = 0.075) +
  geom_point(data=mod_dat, 
             aes(y=pred), fill="grey", 
             size = 6, shape = 21) +
  scale_color_gradient(guide = F,
                       low=other_col, high=r_col) +
  scale_y_continuous(breaks=seq(0,100,25), labels=paste0(seq(0,100,25), "%")) +
  scale_size_continuous(guide = FALSE, range = c(2,6)) +
  labs(title = "Proportions per language family\nand model predictions (world-wide)",
       y = "percentage of forms with /r/", 
       x = "\n") +
  theme_rough +
  theme(plot.margin = unit(c(0.4,0.4,-0.6,0.4),"cm"))
ggsave("graphs/xling_model_preds.png", width=9, height=3.5)

```

