---
title: "Data processing for rough R: Proto-Indo-European forms based on Watkins (2000) for analysis A5"
author: "Márton Sóskuthy"
output: html_document
---

This script takes a raw data file that includes rough/smooth adjectives from the Stadtlander & Murdoch (2000) data set along with etymological data based on Watkins (2000) and creates a clean version for analysis.

Loading packages & data.

```{r}
library(tidyverse)
pie <- read_csv('../raw_data/a5_pie/stadt_rough_hard_words_oed_watkins_2000.csv')
```

Removing words with no roughness rating.

```{r}
pie <- filter(pie, !is.na(RoughnessMean)) %>%
  filter(Word != "wafflish") %>%
  mutate(rough = RoughnessMean >= 0,
         pie.r = str_detect(IE_root_watkins_2000, 'r'))
```

Removing duplicates and NAs.

```{r}
pie <- filter(pie, !is.na(IE_root_watkins_2000)) %>%
  filter(!duplicated(IE_root_watkins_2000))
```

Formatting column names / removing columns that are not needed.

```{r}
pie_final <- pie  %>%
  dplyr::select(-HardnessMean, -MainYear, -EtymonYear,
                -CommentOnOEDFullSense, -DetailedForm_watkins_2000,
                -AlternativeForm) %>%
  dplyr::rename(word="Word",
                roughness="RoughnessMean", etymology="Etymology",
                origin="Origin", root="IE_root_watkins_2000")
```

Saving file.

```{r}
write_csv(pie_final, '../final_data/pie.csv')
```
