---
title: "Data processing for rough R: Proto-Indo-European forms based on Rolf Noyer's database for analysis A6"
author: "Márton Sóskuthy"
output: html_document
---

This script takes a database of Proto-Indo-European roots with reconstructed meanings and extracts relevant surface descriptors based on Stadtlander & Murdoch (2000), adding roughness ratings from this external data set.

Loading packages & data.

```{r}
library(tidyverse)
pie_all <- read_csv('../raw_data/a6_pie_2/all_stems.csv', col_names = FALSE)
eng <- read_csv('../final_data/english_norms.csv')
```

Formatting PIE roots data set + limiting to adjectives only.

```{r}
pie_all <- rename(pie_all, Stem = X1, Description = X2)
adjs <- filter(pie_all, str_detect(Description, 'adj'))
```

The following code chunk is where all the action happens.

```{r}
# create regex to match words in the English norms data
regex <- paste0("[‘ ,;](", str_c(eng$Word, collapse = '|'), ")")

# create PIE roots data set
pie2 <- adjs %>%
  # extract matching surface descriptors from roots data set + the PIE roots
  mutate(surface_descriptor = str_match(adjs$Description, regex)[,2],
         pie_root =  stringr::str_match(adjs$Description, ".*[|][ ]*(.*?)$")[,2]) %>%
  # discard all rows that are not surface descriptors
  filter(!is.na(surface_descriptor)) %>%
  # add in roughness ratings
  left_join(dplyr::select(eng, Word, Rough.M), by=c("surface_descriptor"="Word")) %>%
  mutate(rough = Rough.M >= 0) %>%
  # majority "rough" classification used for duplicates (doesn't seem to change anything!);
  # also, does root contain an /r/?
  group_by(pie_root) %>%
  mutate(rough = as.logical(round(mean(rough))),
         r = str_detect(pie_root, 'r|r̥')) %>%
  ungroup() %>%
  # remove duplicate roots
  filter(!duplicated(pie_root))
```

Saving file.

```{r}
write_csv(pie2, '../final_data/pie2.csv')
```
